How to improve your relationship with your future self.

### 2011-03-21 workflow.Rnw
COntinued editing

\title{A Few Easy Pieces: \\ How to enhance your relationship with your future self.}

@@

An idea is born in a coffee shop, a seminar, a quiet walk. It is a
gray winter in 2011, but the idea dispells February's doldrums and
energizes the student.% Now he knows why he went to graduate school.

This idea inspires a seminar paper in the spring. And a conference
paper arises from the seminar paper in collaboration with another
student in 2012. A dissertation chapter descends from the conference
paper in 2013. Other dissertation chapters take up 2014. A submission
to a journal with the original co-author and a new collaborator
happens in 2015. Revision and resubmission wait until 2017 while
harried editors, reviewers and authors strive to balance teaching,
service, and life. By now, the three lucky collaborators work as
professors in three different universities. In 2018 a child is born
and a paper is published. The United Nations takes an interest in the
paper in 2019 and hosts a conference to discuss implications of the
research. In 2020 a first year graduate student in a coffee shop has
an idea that challenges the analysis in the now famous paper. What
would happen if they had controlled for X? Or run one more MCMC chain?
Or chosen a different likelihood function?  Will the United Nations
(now eager to act based the paper) make a wrong move?

The first author convenes a three way video conference with the other
collaborators during his homeward commute while his flying car is in
auto-drive mode (video chatting during manual driving having been
outlawed in his state in 2017). The group must go back to the
analyses. Which ones?  The ones from 2011? Or 2012? Or 2018? Where are
the files? The next day, one member of the group who has kept some
hard-drives around out of nostalgia (the same guy who still owns
cassette tapes) finds some of the files. Now the checks and
re-analyses should be easy. Right? The student, now professor, is the
same person and should remember the reason for those bits of code (or
at least should remember which series of mouse clicks were used to
produce the numbers for that crucial table --- and those mouse clicks
should not be time consuming to redo exactly as they were done in
2011, or was it 2015?). Right? And, of course, the way Microsoft
Word/Stata/SPSS/R/LISREL stores files and the way that current machines in 2020
read and write them is the same --- since Windows and Mac OS
X have always existed and will always continue to exist more or less
as they currently exist? Right? And the group knows exactly which bit of code
produced which table and which figure, right? And they wrote their
code following Nagler's Maxims \citep{nagler1995coding} and King's
Replication Standard \citep{king1995replication} right?

Right? Wrong? If the collaborators find themselves answering ``wrong'' to many of those questions

If the original analyses were not done via typing in commands in plain
text files, these authors will be in a pickle in 2020.

Just in case one has some doubt about answering these questions, this
piece aims to amplify some of what we already ought to know from
\citet{nagler1995coding} and \citet{king1995replication}, and to
update some of those ideas given current practices, platforms, and
propensities

%%% This data frame will be used throughout the analyses.
%%% Putting it up top, so we can use it in the comments section.

<<read.data,echo=FALSE,results=hide>>=
##  This read.csv line is slightly more complex than needed because of
##  the particular encoding used for characters in the source file.
dem.nations.df<-read.csv(url("http://www.hks.harvard.edu/fs/pnorris/Data/Democracy%20CrossNational%20Data/Democracy%20Crossnational%20Data%20Spring%202009.csv",encoding="ISO_8859-1"),as.is=TRUE)

row.names(dem.nations.df)<-dem.nations.df$Natabrv ##make the row names more helpful

good.df<-na.omit(dem.nations.df[,c("Gini2004","protac2000","Nation","meanpr")])
@.
\section{Type, don't click.}

All tables and graphs should be generated from code, not clicking
(let alone copying and pasting). If I wanted to re-create the figure
you created but with red lines, including a new variable or specification, I
should be able to do so with just a few edits to the code.

Using R, for example, I might specify that the file \Verb+fig1.pdf+
  was produced by the command, where the \Verb+please-plot+ function was
loaded elsewhere.

\begin{Verbatim}[fontsize=\small]
pdf('fig1.pdf')
please-plot(outcome by explanatory)
dev.off()
\end{Verbatim}

@@

\section{Enjoy literate data analysis.}

@@

\section{Enabling revision to prevent clobbering and record history.}

@@

\section{Take advantage of file names to convey information to your future self.}

>>>>

\title{How to improve your relationship with your future self.}

@@

An idea is born in a coffee shop, a seminar, a quiet walk. On this
gray day in 2011, the idea dispells February's doldrums. The student
rushes home, mind racing, the cold ignored.
% Now he knows why he went to graduate school.

This idea inspires a seminar paper in the spring. A conference
paper arises from the seminar paper in collaboration with another
student in 2012. A dissertation chapter descends from the conference
paper in 2013. Other dissertation chapters take up 2014. A submission
to a journal with the original co-author and a new collaborator
happens in 2015. Revision and resubmission wait until 2017 while
harried editors, reviewers and authors strive to balance teaching,
service, and life. By now, the three lucky collaborators work as
professors in three different universities. In 2018 a child is born
and a paper is published. The United Nations takes an interest in the
paper in 2019 and hosts a conference to discuss implications of the
research. In 2020 a first year graduate student in a coffee shop has
an idea that challenges the results in the now famous paper. What
would happen if the authors had controlled for X? Or run one more MCMC chain?
Or chosen a different likelihood function?  Will the United Nations
(now eager to act based the paper) make a wrong move?

The first author convenes a three way video conference with the other
collaborators during his homeward commute after putting his flying car
in auto-drive mode.\footnote{One assumes that video chatting during
  manual driving of flying cars will have been outlawed in his state
  in 2017.} The group must go back to the analyses. Which ones?  The
ones from 2011?  Or 2012? Or 2018? Where are the files? The next day,
one member of the group who has kept some hard-drives around out of
nostalgia finds some of the files.\footnote{This is the same guy who
  still owns cassette tapes and compact discs.} Now the checks and
re-analyses should be easy. Right? The student, now professor, should
remember the reason for those bits of code (or at least should
remember which series of mouse clicks were used to produce the numbers
for that crucial table --- and that mousing should not be time
consuming to redo exactly as they were done in 2011, or was it
2015?). Right? And, of course, the way Microsoft
Word/Stata/SPSS/R/LISREL stores files and the way that current
machines in 2020 read and write them is the same --- since Windows and
Mac OS X have always existed and will always continue to exist more or
less as they currently exist? Right? And the group knows exactly which
bit of code produced which table and which figure, right? And they
wrote their code following Nagler's Maxims \citep{nagler1995coding}
and King's Replication Standard \citep{king1995replication} right?

Right? Wrong? If the collaborators find themselves answering ``wrong''
to many of the questions posed here then reproducing the analyses will
take a lot of time if such reproduction is possible at all. This essay
provides some suggestions for practices which will make such
reproduction occur much more easily and quickly in the event that
famous papers require special scrutiny. Specifically, this piece aims
to amplify some of what we already ought to know from
\citet{nagler1995coding} and \citet{king1995replication}, and to
update some of those ideas given current practices, platforms, and
propensities.

%%% This data frame will be used throughout the analyses.
%%% Putting it up top, so we can use it in the comments section.
<<read.data,echo=FALSE,results=hide>>=
##  This read.csv line is slightly more complex than needed because of
##  the particular encoding used for characters in the source file.
dem.nations.df<-read.csv(url("http://www.hks.harvard.edu/fs/pnorris/Data/Democracy%20CrossNational%20Data/Democracy%20Crossnational%20Data%20Spring%202009.csv",encoding="ISO_8859-1"),as.is=TRUE)
row.names(dem.nations.df)<-dem.nations.df$Natabrv ##make the row names more helpful
good.df<-na.omit(dem.nations.df[,c("Gini2004","protac2000","Nation","meanpr")])
@

\section{Data analysis is computer programming.}

All results (numbers, comparisons, tables, figures) should arise
generated from code, not series of mouse clicks or copying and
pasting. If I wanted to re-create the figure you created but including
a new variable or specification, I should be able to do so with just a
few edits to the code rather than knowledge of how you used your
pointing device in your graphical user interface.

Using R, for example, I might specify that the file \Verb+fig1.pdf+
was produced by the following commands in a file called \Verb+fig1.R+.

<<eval=FALSE>>=
thedata<-read.csv("Data/thedata.csv")
pdf('fig1.pdf')
please-plot(outcome by explanatory using thedata)
dev.off()
@

@@

\section{Data analysis is commnication.}

@@

\section{Version control prevents clobbering and record history.}

@@

\section{File names send messages to your future self.}
